Supplementary material for\\ A Benchmark of Named Entity Recognition Approaches in Historical Documents\\ Application to 19$^{th}$ Century French Directories

### 2022-01-20 src-latex/main-supplementary-material.tex
New paper structure with 1 section for each experiment

% !TeX root = ../main-paper.tex
\section{Dataset}
\label{sec:dataset}

%In this work, we  aim  at  producing  structured  spatio-temporal  data  from the information contained in the XIX\textsuperscript{th} century Parisian trade directories.
%To do this, it is therefore necessary to extract the text from the scans of the directories to be processed and to identify the entries in the various indexing lists used.
%Then, in each of these entries, the named entities they contain have to be extracted to produce structured spatio-temporal data representing the evolution over time of the people and businesses listed in these directories, their descriptions and their locations.
%Both of these extraction tasks require data to train and evaluate the OCR and NER approaches identified as potentially relevant: Tesseract and PERO OCR for the OCR task and Spacy and CamemBERT - pre-trained or fine-tuned - neural models for the NER task.

\subsection{The  XIX\textsuperscript{th} century Parisian trade directories}

The directories to be processed are stored in different libraries in Paris\footnote{\textcolor{green}{Citer les bibliothèques sources?} \joseph{est-ce nécessaire ici ? Est-ce que ça peut être dans une annexe ?}}. They have been scanned independently and with various levels of quality. Moreover, they cover a wide period and have been produced by different publishers and printers. Therefore, their contents, indexes, layouts, methods of printing, etc. may vary a lot from one directory to another (see Figure \ref{fig:directories}).

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/DirectoryExcerpts2.png}}
	  \caption{\label{fig:directories} Examples of directory layouts and contents: 1) Duverneuil et La Tynna 1806 - index by name; 2) Deflandre 1828 - index by activity ; 3) Bottin 1851 –  index by street name}
\end{figure}

For training as testing purposes, our groundtruth dataset must be as representative as possible of the diversity of the available directories. The Table \ref{tab:directories} lists the directories selected, their layout, the structure of their entries, and the number of annotated entries according to the considered index.

For example, the Didot directory of 1854 is organised according to three indexes: by name, by activity, and by street name. The structure of the following entry is characteristic of its index by name: "Batton (D.-A.) \scalerel*{\includegraphics{./images/LH.png}}{H}, professeur au Conservatoire de musique et de déclamation, Saint-Georges, 47." This entry begins with a person's name, followed by the initials of their first name in parentheses.The symbol indicates that this person was awarded the Légion d'Honneur. The following part of this entry represents the person's activity,i.e. their profession, or social status, here "professor at the Conservatory of music and declamation". The street name and number where the person lives or carries out their activity are written at the end of the entry.

These pieces of information constituting the structure of the entries are the basic data that we want to extract, deduplicate and structure in order to build a spatio-temporal database of XIX\textsuperscript{th} century Parisian inhabitants and businesses. With the exception of some potentially wordy activity descriptions, they correspond to named entities. However, as Table \ref{tab:directories} shows, while most entries contain the same types of named entities, their order and the way they are written vary from one directory or index to another. To provide examples of each entry structure, pages from each type have thus been annotated\footnote{In this work, the street name index entries have not been included as they are very different from the others. The strategy to be adopted to process them without loss of quality for the NER model will be the subject of future work.}.

@@

\subsection{Groundtruth data extraction}

To build our ground truth corpus, we annotate 78 pages randomly sampled from each of the major directory series for which we have digitised copies, using a tool developed for our project (see Figure \ref{fig:annotator}). First, the image of a given directory page is loaded. Second, entries are automatically detected and boxes are drawn around them as shown in the left panel of the Graphical User Interface (GUI). Then the text of each entry is extracted by an OCR tool and finally named entities are annotated by a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically. But as they are prone to errors, their results can be corrected manually through the GUI. In the end, the entry bounding boxes, their text, and the named entities are stored in a JSON file.

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/Annot2.png}}
	  \caption{\label{fig:annotator} GUI for corpus annotation \joseph{remove GUI-related details?}}
\end{figure}

The annotated named entities are as follows:
\bertrand{Add a ref to Fig. \ref{fig:annotated-example}}

\begin{itemize}
    \item PER: a person name or a business name, usually referred to as several person names. Firts names, initials, or civility mentions are included. E.g.: "Alibert (Prosper)", "Allamand frères et Hersent", "Heurtemotte (Vve)", etc.
    \item TITLE: an honorary title, either text, glyphs or a combination of the two. E.g: "O. \scalerel*{\includegraphics{./images/LH.png}}{H}" for "Officier de la Légion d'Honneur" or "\scalerel*{\includegraphics{./images/GMLondon.png}}{H}" to represent the Great Medal at the London exhibition.
    \item ACT: the profession or social status of a person summarized in the single concept of activity. E.g.: "horlogerie", "export en tous genres", "Conseiller d'Etat", "propriétaire".
    \item LOC: mostly street names ("r. de la Jussienne"), but may also be neighborhoods ("Marais du Temple") or any indirect spatial references ("au Palais Royal").
    \item CARDINAL: a street number, as part of an address ("16", "5 bis"), or a range of numbers (e.g. "23-25" or "5 à 9").
    \item FT: a geographic entity type, used to give more details on a location, e.g. "boutique", "atelier", "fab." or "dépôt".
\end{itemize}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=\textwidth]{images/examples_labeled_entries.pdf}
    \caption{\label{fig:annotated-example}Two annotated entries showing the different labels used in the corpus.}
\end{figure}
Thus, the extracted corpus contains both the bounding boxes of directory entries, their exact text and the named entities they contain. It can therefore be used to train and improve the entry detection, OCR and NER tools.
% book, page, box, text_ocr, ner_xml (+uid within a page)

\subsection{Assessing the groundtruth quality}
\nathalie{On va utiliser cette sous-section ou pas?}
\joseph{Pas vraiment car on n'a pas eu le temps de faire ça, mais on peut parler de l'étage d'extraction d'entrées qui est validé manuellement, et des quelques règles de simplification du charset pour homogénéiser entre les annotateurs / pallier des ambiguïtés dans les règles d'annotation (qu'on pourrait encore améliorer)}
Our dataset is thus made of excerpts TODO interesting points to detail:

\begin{itemize}
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we consider and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assess clean text's quality)
    \item Measure inter-annotator agreement ? 
    \item Explain charset projection
    \item \dots
\end{itemize}

>>>>

\documentclass[runningheads,svgnames]{llncs}
%
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{orcidlink}
\usepackage{subcaption} % Addition
\usepackage{multirow} % Addition
\usepackage{tabularx} % Addition
\usepackage{siunitx} % Addition
\usepackage{scalerel}
\usepackage[inline,shortlabels]{enumitem}

% \usetikzlibrary{shadings,decorations,decorations.pathreplacing}
\renewcommand\UrlFont{\color{blue}\rmfamily}
% FIXME proper hyperref setup

\newcommand{\mcite}[1]{\text{\cite{#1}}}
\newcommand{\mref}[1]{\text{\ref{#1}}}

% Customize table rendering
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.4}

\begin{document}
%
\title{%
Supplementary material for\\
A Benchmark of Named Entity Recognition Approaches in Historical Documents\\
Application to 19$^{th}$ Century French Directories%
}
%
\titlerunning{Supplementary material}
%
\author{%
N. Abadie\inst{1}\orcidlink{0000-0001-8741-2398} \and
E. Carlinet\inst{2}\orcidlink{0000-0001-5737-5266} \and
J. Chazalon\inst{2}\orcidlink{0000-0002-3757-074X} \and
B. Duménieu\inst{3}\orcidlink{0000-0002-2517-2058}\\
{\footnotesize \emph{all authors contributed equally}}%
}
%
\authorrunning{N. Abadie et al.}
%
\institute{%
% 1
LASTIG, Univ. Gustave Eiffel, IGN-ENSG, F-94160 Saint-Mandé, France\\
\email{nathalie-f.abadie@ign.fr}%\\
% \url{https://www.umr-lastig.fr/nathalie-abadie/}%
\and
% 2
EPITA Research \& Development Laboratory (LRDE), Le Kremlin-Bicêtre, France\\
\email{\{edwin.carlinet,joseph.chazalon\}@lrde.epita.fr}%\\
%\url{https://lrde.epita.fr}
\and
% 3
CRH-EHESS, Paris, France\\
\email{bertrand.dumenieu@ehess.fr}%\\
% \url{http://crh.ehess.fr/index.php?5206}\\
}
%
\maketitle
%
\begin{abstract}
This document contains supplementary material for the paper ``A Benchmark of Named Entity Recognition Approaches in Historical Documents ---
Application to 19$^{th}$ Century French Directories''.
\end{abstract}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% #####     ##     #####    ##     ####   ######   #####
% #    #   #  #      #     #  #   #       #          #
% #    #  #    #     #    #    #   ####   #####      #
% #    #  ######     #    ######       #  #          #
% #    #  #    #     #    #    #  #    #  #          #
% #####   #    #     #    #    #   ####   ######     #
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dataset}

\subsection{Content}

@@

\subsubsection{Annotation tools}
To build our ground truth corpus, we annotate 78 pages randomly sampled from each of the major directory series for which we have digitised copies, using a tool developed for our project (see Figure \ref{fig:annotator}). First, the image of a given directory page is loaded. Second, entries are automatically detected and boxes are drawn around them as shown in the left panel of the Graphical User Interface (GUI). Then the text of each entry is extracted by an OCR tool and finally named entities are annotated by a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically. But as they are prone to errors, their results can be corrected manually through the GUI. In the end, the entry bounding boxes, their text, and the named entities are stored in a JSON file.

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/Annot2.png}}
	  \caption{\label{fig:annotator} GUI for corpus annotation}
\end{figure}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=\textwidth]{images/examples_labeled_entries.pdf}
    \caption{\label{fig:annotated-example}Two annotated entries showing the different labels used in the corpus.}
\end{figure}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% #####   ######   ####   #    #  #        #####   ####
% #    #  #       #       #    #  #          #    #
% #    #  #####    ####   #    #  #          #     ####
% #####   #            #  #    #  #          #         #
% #   #   #       #    #  #    #  #          #    #    #
% #    #  ######   ####    ####   ######     #     ####
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

% ---- Bibliography ----
% \bibliographystyle{splncs04}
% \bibliography{ref,ref-supp}
\end{document}

### 2022-01-20 src-latex/parts/20-related-work.tex
Add annotated example
Shrink intro
New paper structure with 1 section for each experiment

% !TeX root = ../main-paper.tex
\section{OCR and NER on historical texts}

@@

\nathalie{à compléter/retoucher?}

@@

\subsection{Named Entity Recognition on historical texts}

\cite{Labusch2020NamedED} pretrained an off-the-shelf multilingual BERT-based NER model on more than 2 000 000 pages of OCRed historical texts in German, French, and English. They fine-tuned it on historical texts NER ground truth, also in these three languages and tested it on OCRed historical texts in each language.
Eventually, a clear decrease in NER performance on OCRed texts is noted, especially for the English texts for which the OCR is of poorer quality.

@@

>>>>

% !TeX root = ../main-paper.tex
\section{OCR and NER on historical texts}
\label{sec:related-work}

@@

\begin{itemize}
    \item Tesseract 4.1.1 (newest version: 5, released Nov. 2021)
    \item Pero OCR version from git repo, master branch updated on Sep 15, 2021 % 8b20f29    
    \item other?  % kraken if times allows
\end{itemize}
Pero OCR uses ParseNet \joseph{TODO ref} as internal layout parser for line detection.
Pero OCR uses an LSTM engine \joseph{TODO check + ref} for text line recognition.
Pero OCR generally works very well, as long as the bounding boxes of the regions to recognize are not too tightly adjusted.
Trained on newspapers in European languages by its development team.
Can output results in Latin, Greek and Cyrillic scripts, as well as some commonly-used typographic symbols.

% PERO refs
% O Kodym, M Hradiš: Page Layout Analysis System for Unconstrained Historic Documents. ICDAR, 2021.
% M Kišš, K Beneš, M Hradiš: AT-ST: Self-Training Adaptation Strategy for OCR in Domains with Limited Transcriptions. ICDAR, 2021.
% J Kohút, M Hradiš: TS-Net: OCR Trained to Switch Between Text Transcription Styles. ICDAR, 2021.

@@

\subsection{Named Entity Recognition on historical texts}
\label{subsection:stoa-ner-on-historical-texts}
\cite{Labusch2020NamedED} pretrained an off-the-shelf multilingual BERT-based NER model on more than 2 000 000 pages of OCRed historical texts in German, French, and English. They fine-tuned it on historical texts NER ground truth, also in these three languages and tested it on OCRed historical texts in each language.
Eventually, a clear decrease in NER performance on OCRed texts is noted, especially for the English texts for which the OCR is of poorer quality.

@@

\subsection{== À réintégrer depuis ancienne section 4 ==}
We select two deep-learning-based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.

\subsubsection{spacy}
Spacy is a software library that offers NLP components assembled in modular pipelines specialised by language.
Although BERT is available in the latest version of SpaCy (v3), the pipeline for French does not provide a NER layer at the time of our experiments (jan. 2022).
Hence, we rely on SpaCy's ad hoc pipeline trained on French corpora and capable of named entity recognition.
% deep-sequoia and wikiner-fr, capable of named entity recognition.\textit{fr\_core\_news\_lg}\footnote{https://spacy.io/models/fr} trained two corpora in French: deep-sequoia and wikiner-fr, capable of named entity recognition.
The global architecture of these pipelines have not been yet published but are explained by the developers on their website.
Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
% \bertrand{Drop the next sentences if we need space}
% In 2018~\cite{won2018} evaluated SpaCy's NER ability to detect place names in five corpora of ancient letters written in English.
% They measured an average F1 score of 0.57.
% The SpaCy developers claim an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.
%For our experiments, the French model \textit{fr\_core\_news\_lg} is fine-tuned using our ground truth corpus.

% Other possibilities:
%1. Traditional ML based:
%    Conditional Random Fields (CRF) - https://pypi.org/project/sklearn-crfsuite/
%    Maximum-entropy Markov model

%2. Neural Networks based:
%    LSTMs, bi-LSTM - https://github.com/flairNLP/flair
%    CNNs (SpaCy uses CNN based architecture)
%    Transformers (Spacy has recently launched it) - %https://spacy.io/universe/project/spacy-transformers

\subsubsection{Transformers / Bert / Hugging face}
Language model based on Transformers, like BERT, have become a new paradigm for NER\cite{li2020}. 
The learned embeddings can be used as distributed representations for input instead of traditional embeddings like Google Word2Vec, and they can be further fine-tuned for NER by adding an additional output layer. 
They can also be pretrained in an unsupervised way on historical texts for domain adaptation.
As the directories are written in French, we chose the language model CamemBERT \cite{martin-etal-2020-camembert}, a Transformer model trained on a French corpus.
\joseph{plutôt tourner ça pour dire qu'il existe des modèles FR / qu'on est limités par la dispo de modèles FR pré-entraînés ?}

\subsection{Pipeline summary}

\Cref{fig.protocol} depicts the evaluation protocol used to assess the OCR and NER systems.

\begin{figure}[tb]
    \includegraphics[width=\linewidth]{figs/protocol.pdf}
    \caption{Scheme of the evaluation protocol. \joseph{Missing "Ground truth" for NER stage?}}
    \label{fig.protocol}
    \end{figure}

### 2022-01-20 src-latex/parts/30-dataset.tex
Shrink intro
New paper structure with 1 section for each experiment

% !TeX root = ../main-paper.tex
\section{Dataset}

@@

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/DirectoryExcerpts2.png}}
	  \caption{\label{fig:directories} Examples of directory layouts and contents: 1) Duverneuil et La Tynna 1806 - index by name; 2) Deflandre 1828 - index by activity ; 3) Bottin 1851 –  index by street name}
\end{figure}

@@

\begin{table}[h!]
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Directory publisher or name}} & \multicolumn{1}{c|}{\textbf{Year}} & \multicolumn{1}{c|}{\textbf{\# Col}} & \multicolumn{1}{c|}{\textbf{Index}} & \multicolumn{1}{c|}{\textbf{Entry structure}} & \multicolumn{1}{c|}{\textbf{\# Ann}}                                                                                                      
\\ \hline
Favre \& Duchesne & 1798 & 1 & by activity & {N [A], SN, SNUM - SEC.} & 233
\\ \hline
Duverneuil et La Tynna & 1801 & 1 & \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & {N, SN, SNUM. SEC.} & \begin{tabular}[c]{@{}l@{}} 136 \\ 103\end{tabular}
\\ \hline
{Notables communaux de la Seine} & 1801 & 1 & by name & {N, [A,] SN[, SNUM].} & 92
\\ \hline
Duverneuil et La Tynna & 1805 &	2 &	\begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}}N [(F)], SN, SNUM. - SEC.\\ N [(C)], [(A),] SN, SNUM. - SEC.\end{tabular} & \begin{tabular}[c]{@{}l@{}} 155 \\ 221 \end{tabular}
\\ \hline
Duverneuil et La Tynna & 1806 & 2 & \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}} N [(F)], SN, SNUM.\\ N [(C)], SN, SNUM.[ - SEC.]\end{tabular}. & \begin{tabular}[c]{@{}l@{}} 186\\ 0\end{tabular}
\\ \hline
La Tynna & 1813	& 2	& \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}} N,[A,] SN[, SNUM].\\ N, [(T)] A, SN[, SNUM] [, D].\end{tabular} & \begin{tabular}[c]{@{}l@{}} 135 \\ 177 \end{tabular}
\\ \hline
Panckoucke Commerces (Dulac) & 1820 & 1 & by name & N [(F)], A, [(T),] SN, SNUM[, P]. & 0
\\ \hline
Panckoucke Habitants (Dulac) & 1820	& 1 & by name & N [(T)], [A,] SN, SNUM [, D]. & 0
\\ \hline
Bottin (serie 1: 1819-1838) & 1820 & 2	& \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}} N [(F \textbar C)][T], [P], SN, SNUM.\\N [(F \textbar C)][T], A, SN, SNUM. CR\end{tabular} & \begin{tabular}[c]{@{}l@{}} 128 \\ 151 \end{tabular}
\\ \hline
Bottin (serie 1: 1819-1838) & 1827 & 2	& \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}}N [(F \textbar C)][T], [P], SN, SNUM.\\ N [(F \textbar C)][T], A, SN, SNUM.\end{tabular} & \begin{tabular}[c]{@{}l@{}} 228 \\ 285 \end{tabular}
\\ \hline
Deflandre & 1828 & 2 & \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}} N [(F \textbar C)][T], [P], SN, SNUM. \\N [(F \textbar C)][T], A, SN, SNUM.\end{tabular} & \begin{tabular}[c]{@{}l@{}}115\\ 229\end{tabular}
\\ \hline
Deflandre & 1829 & 2 & \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}}N [(F \textbar C)][T], [P], SN, SNUM.\\ N [(F \textbar C)][T], A, SN, SNUM.\end{tabular} & \begin{tabular}[c]{@{}l@{}}177\\ 236\end{tabular}
\\ \hline 
Bottin (serie 1: 1819-1838) & 1837 & 2	& \begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}}N [(F \textbar C)][T], [P], SN, SNUM.\\ N [(F \textbar C)][T], [A], SN, SNUM.\end{tabular} & \begin{tabular}[c]{@{}l@{}}158\\ 557\end{tabular}
\\ \hline
Cambon - Almanach général & 1841 & 2 &	\begin{tabular}[c]{@{}l@{}}by activity\\ by name\end{tabular} & \begin{tabular}[c]{@{}l@{}}N [(F \textbar C)][T], [P], SN, SNUM.\\ N [(F \textbar C)][T], A, SN, SNUM.\end{tabular} & \begin{tabular}[c]{@{}l@{}}182\\ 486\end{tabular}
\\ \hline
Didot & 1841a &	\begin{tabular}[c]{@{}l@{}}3\\ 4\end{tabular} &	\begin{tabular}[c]{@{}l@{}}by activity \\ by name \end{tabular} & 	\begin{tabular}[c]{@{}l@{}}N [(F \textbar C)][T], [P], SN, SNUM. \\ N [(F \textbar C)][T], [A], SN, SNUM. \end{tabular} & \begin{tabular}[c]{@{}l@{}}246\\ 843\end{tabular}
\\ \hline
Didot & 1851a & \begin{tabular}[c]{@{}l@{}l@{}}4\\ 4\\ 5\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}}by activity\\ by name\\ by street name\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}}N [(F \textbar C)][T], [P], SN, SNUM.\\ N [(F \textbar C)][T], [A], SN, SNUM.\\SNUM N [(F \textbar C)][T], [A].\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}}309\\ 961\\ 0\end{tabular}
\\ \hline
Didot & 1854a & \begin{tabular}[c]{@{}l@{}l@{}}4\\ 3\\ 5\end{tabular} &	\begin{tabular}[c]{@{}l@{}l@{}} by activity \\ by name \\ by street name\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}} N [(F \textbar C)], [P], SN, SNUM [,T].\\ N [(F \textbar C)][T], [A], SN, SNUM.\\ SNUM N [(F \textbar C)][T], [A].\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}}106\\ 362\\ 0\end{tabular}
\\ \hline
Bottin (serie 3: 1854-1856) & 1854a & \begin{tabular}[c]{@{}l@{}l@{}}3\\ 2\\ 4\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}} by activity \\ by name \\ by street name\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}}N [(F \textbar C)], [P], SN, SNUM [,T].\\ N [(F \textbar C)][T], [A], SN, SNUM.\\ SNUM N [(F \textbar C)], [A] [,T].\end{tabular} & \begin{tabular}[c]{@{}l@{}l@{}}253\\ 633\\ 0\end{tabular}
\\ \hline
Didot-Bottin & 1860a	& 3	& by name & N [(F \textbar C)][T], [A], SN, SNUM. & 378
\\ \hline
Didot-Bottin & 1861a	& 3	& by name & N [(F \textbar C)][T], [A], SN, SNUM. & 402
\\ \hline
\end{tabular}%
}
\caption{The directories used to build our dataset. Letters next to year of publication are used to identify the volume of the directory. \# Col stands for "Number of columns", and \# Ann stands for "Number of annotated entries". Entry structure is described as follows: N - entry name (person or business); F - first name; T - honorary title; C - civility; A - activity; P - precision; SN - street name; SNUM - street number; SEC - section; D - district; CR - cross-reference; bracketed elements are optional.}
\label{tab:directories}
\end{table}

\subsection{Groundtruth data extraction}

To build our ground truth corpus, we annotate 78 pages randomly sampled from each of the major directory series for which we have digitised copies, using a tool developed for our project (see Figure \ref{fig:annotator}). First, the image of a given directory page is loaded. Second, entries are automatically detected and boxes are drawn around them as shown in the left panel of the Graphical User Interface (GUI). Then the text of each entry is extracted by an OCR tool and finally named entities are annotated by a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically. But as they are prone to errors, their results can be corrected manually through the GUI. In the end, the entry bounding boxes, their text, and the named entities are stored in a JSON file.

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/Annot2.png}}
	  \caption{\label{fig:annotator} GUI for corpus annotation \joseph{remove GUI-related details?}}
\end{figure}

The annotated named entities are as follows:
\bertrand{Add a ref to Fig. \ref{fig:annotated-example}}

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=\textwidth]{images/examples_labeled_entries.pdf}
    \caption{\label{fig:annotated-example}Two annotated entries showing the different labels used in the corpus.}
\end{figure}
Thus, the extracted corpus contains both the bounding boxes of directory entries, their exact text and the named entities they contain. It can therefore be used to train and improve the entry detection, OCR and NER tools.
% book, page, box, text_ocr, ner_xml (+uid within a page)

\subsection{Assessing the groundtruth quality}
\nathalie{On va utiliser cette sous-section ou pas?}
\joseph{Pas vraiment car on n'a pas eu le temps de faire ça, mais on peut parler de l'étage d'extraction d'entrées qui est validé manuellement, et des quelques règles de simplification du charset pour homogénéiser entre les annotateurs / pallier des ambiguïtés dans les règles d'annotation (qu'on pourrait encore améliorer)}
Our dataset is thus made of excerpts TODO interesting points to detail:

\begin{itemize}
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we consider and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assess clean text's quality)
    \item Measure inter-annotator agreement ? 
    \item Explain charset projection
    \item \dots
\end{itemize}

>>>>

% !TeX root = ../main-paper.tex
\section{Dataset}
\label{sec:dataset}

@@

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/DirectoryExcerpts2.png}}
	  \caption{\label{fig:directories} Examples of directory layouts and contents: 1) Duverneuil et La Tynna 1806 - index by name; 2) Deflandre 1828 - index by activity ; 3) Bottin 1851 - index by street name}
\end{figure}

@@

\subsection{Groundtruth data extraction}
\joseph{TODO: separate entry detection, OCR and NER stages explicitely}

First, the layout extraction and entry segmentation of the page are performed with a semi-automated system and
checked by a human. Afterward, an OCR system runs on the thumbnails of each segmented entry to extract its text. An
entry might span over multiple text lines but is always a single block. Thus, the most adapted mode is chosen when
the OCR system allows for the detection mode (e.g. the \emph{block} mode for \emph{Tesseract}).

OCR: some basic charset normalization is performed after human annotation:
- remove annotation codes for glyphs which have no character in unicode
- alleviate some ambiguities in annotation rules (n° -> no, --- -> -, homogenize star and hand symbols allowed)

The annotated named entities are as follows:

@@

Thus, the extracted corpus contains both the bounding boxes of directory entries, their exact text and the named entities they contain. It can therefore be used to train and improve the entry detection, OCR and NER tools.
% book, page, box, text_ocr, ner_xml (+uid within a page)

\subsection{Summary}
\joseph{TODO intégrer dans les sections avant ?}

\begin{verbatim}
[('<LOC>', 9709),
('<PER>', 8788),
('<CARDINAL>', 8747),
('<ACT>', 6472),
('<TITRE>', 483),
('<FT>', 43)]
\end{verbatim}

 valid ocr: 8765 entries

 valid ner (ref \& pero \& tess): 8341

\joseph{parler de l'étage d'extraction d'entrées qui est validé manuellement, et des quelques règles de simplification du charset pour homogénéiser entre les annotateurs / pallier des ambiguïtés dans les règles d'annotation (qu'on pourrait encore améliorer)}

TODO a word about the charset, and existence of glyphs unknown to the OCR

TODO average entry len (in char)

TODO average number of tags/entities per entry

TODO lists actual dataset content: set of entries. For each entry:
\begin{itemize*}
    \item ref to original book, page, region
    \item human transcription (OCR ref)
    \item human NER (NER ref)
    \item OCR produced by Pero OCR, Tesseract v4, Kraken (precise model used)
\end{itemize*}

### 2022-01-20 src-latex/parts/40-ner-approaches-BACKUP.tex
New paper structure with 1 section for each experiment

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select two deep-learning-based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
We also create an additional CamemBERT model pretrained on a collection of unannotated directory entries extracted with Pero-OCR.
The NER layer of each three models is then fine-tuned on historical directories as detailed in\cref{subsection:experiment-1-setup} and\cref{subsection:experiment-2-setup}.

@@

\subsubsection{SpaCy NER pipeline~[SpaCy NER]}
Spacy is a software library that offers NLP components assembled in modular pipelines specialised by language.
Although BERT is available in the latest version of SpaCy (v3), the pipeline for French does not provide a NER layer at the time of our experiments (jan. 2022).
Hence we rely on SpaCy's ad hoc pipelines trained on French corpora and capable of named entity recognition.
% deep-sequoia and wikiner-fr, capable of named entity recognition.\textit{fr\_core\_news\_lg}\footnote{https://spacy.io/models/fr} trained two corpora in French: deep-sequoia and wikiner-fr, capable of named entity recognition.
The global architecture of these pipelines have not been yet published but are explained by the developers on their website.
Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
\bertrand{Drop the next sentences if we need space}
In 2018~\cite{won2018} evaluated SpaCy's NER ability to detect place names in five corpora of ancient letters written in English.
They measured an average F1 score of 0.57.
The SpaCy developers claim an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.
%For our experiments, the French model \textit{fr\_core\_news\_lg} is fine-tuned using our ground truth corpus.

>>>>

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select two deep-learning-based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
We also create an additional CamemBERT model pretrained on a collection of unannotated directory entries extracted with Pero-OCR.
The NER layer of each three models is then fine-tuned on historical directories as detailed in \cref{subsection:experiment-1-setup} and \cref{subsection:experiment-2-setup}.

@@

\subsubsection{SpaCy NER pipeline~[SpaCy NER]}
Spacy is a software library that offers NLP components assembled in modular pipelines specialised by language.
Although BERT is available in the latest version of SpaCy (v3), the pipeline for French does not provide a NER layer at the time of our experiments (jan. 2022).
Hence, we rely on SpaCy's ad hoc pipeline trained on French corpora and capable of named entity recognition.
% deep-sequoia and wikiner-fr, capable of named entity recognition.\textit{fr\_core\_news\_lg}\footnote{https://spacy.io/models/fr} trained two corpora in French: deep-sequoia and wikiner-fr, capable of named entity recognition.
The global architecture of these pipelines have not been yet published but are explained by the developers on their website.
Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
\bertrand{Drop the next sentences if we need space}
In 2018~\cite{won2018} evaluated SpaCy's NER ability to detect place names in five corpora of ancient letters written in English.
They measured an average F1 score of 0.57.
The SpaCy developers claim an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.
%For our experiments, the French model \textit{fr\_core\_news\_lg} is fine-tuned using our ground truth corpus.
